---
title: "Project_5310"
author: "Malik"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Red To Green Move
WE came up with a day trading strategy which We want to test. strategy is simple. Any FAANG stock that opens up few percent below the previous closing price, will rally back to the previous closing price. FAANG is an acronym that refers to five prominent publicly traded American technology companies: Facebook (FB), Amazon (AMZN), Apple (AAPL), Netflix (NFLX); and Alphabet (GOOG) (formerly known as Google).
for example, if Apple's closing price on monday was 100 and Tuesday opening price is 99, which is 1% below the monday's closing price, according to our strategy, the probability of stock going back to 100, at least for a moment, is very high. since its a day trade, we buy at the open and sell at the previous closing price.For this analysis, we are going to use change percentage, volume, and 3 techincal indicators Moving averages, Bollinger bands, and Relative strength index. We will explain each of the techincal indicators as we move forward with our analysis. We are going to use historic stock price data.  

```{r}
library(dplyr)
library(quantmod) ## using getSymbol function from quantmod library to get the stock price from www.yahoo.com. this can get us latest price as well.
library(data.table)

facebook <- getSymbols("FB", src = "yahoo", from = "2013-01-01", to = "2020-11-17", auto.assign = FALSE)

head(facebook)
```


#### Lets quickly visualize our data. We are going to use candlestick chart for the month of Jan and Feb 2020. 
```{r}
candleChart(facebook, TA=c(addVo()), subset = '2020-01::2020-02', theme = "white") ## candleChart is a function of quantmod library .TA allows for the inclusion of a variety of chart overlays and tecnical indicators.. In this particular Chart, we are adding volume at the bottom. 

```

**Each candlestick represent single day.length of the candle represents high and the low of the day. candle will be green if it closes above the open price and orange or red if it closes below the open price of the same day. At the bottom of the chart, we have volume in millions. here is a link to understand candlestick in details https://www.investopedia.com/trading/candlestick-charting-what-is-it/ **

## Moving Averages
#### Moving averages are one of the most common techincal indicators in stocks. It helps analysts identify trends in price fluctuations. Lets add Moving averages to our Dataframe and Visualize it.

```{r}

# adding 200 and 50 day moving averages using SMA function. 
facebook <- cbind(facebook, round(SMA(facebook$FB.Adjusted, 50),2)) 
facebook <- cbind(facebook,round(SMA(facebook$FB.Adjusted, 200),2))

candleChart(facebook, TA=c(addVo(),addSMA(n=200,col = 'blue'),addSMA(n=50,col = 'red')),  subset = '2020-01::2020-10', theme = "white")
```

**In this chart we are using a longer period of time to see how moving averages lines behave over a time period. Red line is a 50 day MA and blue line is 200DMA. The 200-day moving average is considered especially significant in stock trading. As long as the 50-day moving average of a stock price remains above the 200-day moving average, the stock is generally thought to be in a upward trend. A crossover to the downside of the 200-day moving average is interpreted as downward trend. here is the link to learn more about Moving averages https://www.investopedia.com/ask/answers/122414/what-are-most-common-periods-used-creating-moving-average-ma-lines.asp **


## Relative Strength Index (RSI)

#### The relative strength index (RSI) is a momentum indicator used in technical analysis that measures the magnitude of recent price changes to evaluate overbought or oversold conditions in the price of a stock or other asset. RSI values are from 0 to 100 Lets add Relative strength Indnex using a RSI function and viusalize it.

```{r}
facebook <- cbind(facebook, round(RSI(facebook$FB.Adjusted),2))

candleChart(facebook, TA=c(addVo(),addSMA(n=200,col = 'blue'),addSMA(n=50,col = 'red'), addRSI(n=14)),  subset = '2020-01::2020-10', theme = "white")
```

**Technical traders use the centerline to show shifts in the trend. If RSI is above 50, momentum is considered up and traders can look for opportunities to buy the market. A drop below 50 would indicate the development of a new bearish market trend. here is a link to learn more about RSI indicator https://www.investopedia.com/articles/active-trading/042114/overbought-or-oversold-use-relative-strength-index-find-out.asp**

## Bollingers Bands

#### Bollinger Bands are a technical indicator that forms a channel around the price movements of an asset.Bollinger bands have three lines, an upper, middle and lower. The middle line is a moving average of prices,The upper and lower bands are drawn on either side of the moving average. The distance between the upper and lower band is determined by standard deviations. The trader determines how many standard deviations they want the indicator set at, although many use two standard deviations from the average.lets add Bolliner bands and visualize it.

```{r}
facebook <- cbind(facebook,round(BBands(HLC(facebook)),2))

candleChart(facebook, TA=c(addVo(),addSMA(n=200,col = 'blue'),addSMA(n=50,col = 'red'), addRSI(n=14),addBBands()),  subset = '2020-01::2020-10', theme = "white")
```

**Red doted lines around the candlesticks are upper and lower bounds. gray dotted line is the average. by defualt, average is 20SMA and bounds  are  average +-  2 standard deviation. Traders tend to buy over the average and sell below.**


## Data cleaning and Transforming 

#### In order to run our logistic regression, we need to find the the percentage difference of previous close and the next day open. here is how it works, ((open - previous close)/open) * 100.

```{r}

# rename the column names so we can see it better.
colnames(facebook) <- c("Open", "High", "Low", "Close", "Volume", "Adjusted", "200DMA", "50DMA", "RSI","Lower-bound","Mavg","Upper-bound", "PctB")

head(facebook)

## because over 200 day moving average column needed previous 200 days worth of data to calculate 200 DMA, it will show nan values in first 200 days. we can delete those rows so we dont have any NA values.

facebook <- facebook[-(1:200), drop = FALSE]

head(facebook)

## check data class

class(facebook)
# convert variable data type to dataframe.
fb <- as.data.frame(facebook)
# add column name for the date column
fb <- cbind(date=row.names(fb),fb)
## Add a new column to dataframe 
fb <-  mutate(fb, Percent_Difference = ((lead(fb$Open) - fb$Adjusted)/lead(fb$Open)*100))
## round up your percentage 
fb$Percent_Difference <- round(fb$Percent_Difference,2)
## move down your differences so we can allign with the correct date.
fb$Percent_Difference <- shift(fb$Percent_Difference,1)

```


#### last and a very important step to do is to create a new column that calculate winning trade and lossing trade. Remember, our theory is, if the opening price is lower than the previous closing price by a small percentage, it will rally back to the previous closing price for at least a moment. For that, we compare open price to high price. if the high price is greater then open price, it is a wining trade.

```{r}
fb["Win/Loss"] <- ifelse(fb$High >  shift(fb$Close) & fb$Percent_Difference < 0 ,1, 0)

 ## since we are only concern with the days where stock opened lower than previous close, we will use filter to get negative values in percent_Difference column.

filter_data <- filter(fb, Percent_Difference < 0)
head(filter_data,20)

```



```{r}
## creating a function that can get any stock price from a given time period and create columns that we need for analysis.

FAANG <- function(ticker){
  stock_data <- getSymbols(ticker, src = "yahoo", from = "2013-01-01", to = "2020-11-17", auto.assign = FALSE)
  colnames(stock_data) <- c("Open", "High", "Low", "Close", "Volume", "Adjusted")
  stock_data <- cbind(stock_data, round(SMA(stock_data$Adjusted, 50),2)) 
  stock_data <- cbind(stock_data, round(SMA(stock_data$Adjusted, 200),2))
  stock_data <- cbind(stock_data, round(RSI(stock_data$Adjusted),2))
  stock_data <- cbind(stock_data, round(BBands(HLC(stock_data)),2))
  colnames(stock_data) <- c("Open", "High", "Low", "Close", "Volume", "Adjusted", "200DMA", "50DMA",   "RSI","Lower-bound","Mavg","Upper-bound", "PctB")
  stock_data <-  stock_data[-(1:200), drop = FALSE]
  stock_data <- as.data.frame(stock_data)
  stock_data <- cbind(date=row.names(stock_data),stock_data)
  stock_data <-  mutate(stock_data, Percent_Difference = ((lead(stock_data$Open) - stock_data$Adjusted)/lead(stock_data$Open)*100))
  stock_data$Percent_Difference <- round(stock_data$Percent_Difference,2)
  stock_data$Percent_Difference <- shift(stock_data$Percent_Difference,1)
  stock_data["Win/Loss"] <- ifelse(stock_data$High >  shift(stock_data$Close) & stock_data$Percent_Difference < 0 ,1, 0)
  stock_data <- filter(stock_data, Percent_Difference < 0)
  return(stock_data)
}

## since we are interested in FAANG stocks, we will only use these stock 
test <- FAANG("FB")
Amazon <- FAANG("AMZN")
Apple <- FAANG("AAPL")
Netflix <- FAANG("NFLX")
Google <- FAANG("GOOG")

head(test)

```



